The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-08-28 manuscript/rebuttal.Rmd

RESPONSE: We appreciate reviewer 2's positive feedback, insights and thoughtful suggestions.

### R2 Point 1

> 1. Inclusion of Case Studies or Examples: To strengthen your argument, consider incorporating case
> studies or practical examples that demonstrate the effectiveness of your recommendations.
> Highlighting specific instances where these best practices have enhanced reproducibility or solved
> particular issues could substantiate your points more effectively.

RESPONSE: To demonstrate the importance of computational reprodcibility, we elaborate on the
"forensic bioinformatics" case study of Baggerly and Coombes (2010) @Baggerly2010-nr critique of Potti
et al (2006) @Potti2006-yk in the introduction:

@@

## Reviewer 3

### R3 Point 1

> The paper is mostly well written and addresses an important topic: computational reproducibility and
> tools to achieve it. The authors describe five "pillars" under which various tools and techniques can
> be categorized. The paper is pretty comprehensive in what it covers and has useful supplementary
> material. My main issue with the paper is that it reiterates many things that have been covered in
> other papers. The authors cite 11 such papers. Many of the topics covered in submitted paper have
> already been covered well in other papers. I am most familiar with reference 13, which covers many of
> the same topics, although this paper goes into more detail on many issues. The authors could do more
> to differentiate their paper from previous ones and perhaps remove some topics that are already
> covered well elsewhere.

RESPONSE: We thank reviewer 3 for their careful assessment of our article and appreciate the suggestions
which have improved the manuscript.
Indeed, this article does reiterate a great deal of what has already been said, it is a review after
all.
We note that each of those recommendation articles mentioned have different emphases.
For example, some of them don't speak to the benefits of literate programming,
while others do not mention the importance of documentation.
Some are focused on specific concepts like infrastructure, containerisation or
continuous analysis.
Some are proof of principle workflows, which present a rigid combination of tools.
Here, we describe the ways computational researchers can make their own work more reproducible.
We are assuming that the readers are not advanced analysts, so some topics like pipelining systems
are covered in less detail and we refer the reader to those relevant reviews.
What does make this article unique is the five pillars infographic, and specific combination of
recommendations that incorporates new material (eg: Software Heritage, Guix, Continuous Analysis).
This is useful, because bioinformatics analysts can use the infographic to benchmark their projects
and identify areas where changes are needed.
Our article also provides a handy list of recommended tutorials as a supplementary resource.

@@

### R3 Point 7

> 7. Page 5, Line 15, type-o = "authors provided along the"

RESPONSE: The passage in question has been amended:

> In this work, authors provided alongside the traditional research paper a Jupyter notebook ...

@@

> While deidentification of data has been widely used to safeguard participant privacy, this may not be
> suitable for genomics data due to the availability of existing public genetic data sets that heighten
> reidentification risks [@Heeney2011-ri].
> For example, the surnames of some male participants can be inferred based on the Y chromosome
> variants seen in deidentified genomic data together with public geneology web sites [@Gymrek2013-fz].
> To foster the responsible reuse of sensitive genomic and health data, the Global Alliance for
> Genomics and Health (GA4GH) initiative has proposed strategies, technical standards and policy
> frameworks designed to protect personal data in a way that perserves reproducibility [@Rehm2021-kt;@Thorogood2021-kt].

### R3 Point 19

> 19. Figures 2 and 4 are very similar to figures used in reference 13.

RESPONSE: Figure 4 is adapted from the diagram on the Docker official website, which was originally published in
2015 or earlier (https://web.archive.org/web/20151023003338/https://www.docker.com/what-docker).
We have cited the original source.

@@

We do not believe any changes to Figure 2 or 4 are necessary.

### R3 Point 20

> 20. Figure 5: I don't think it's really necessary to resummarize the FAIR principles.
> People can go to the source article for that.

RESPONSE: We acknowledge that it can sometimes be tiresome to reiterate things like the FAIR principles.
However summarising the key points succinctly will be in the interests of readers with limited time.
We note that the original description of the FAIR principles by Wilkinson and colleagues is ~5,000 words,
while this entire manuscript is ~7,000. We note that other reviewers did not insist on removing this
section.

@@

### R3 Point 21

> 21. Page 11, line 13: I disagree that repositories like GEO and SRA are FAIR.
There are lots of problems with FAIRness in these repositories.

RESPONSE: While there may be problems with FAIRness with NCBI GEO and SRA, these repositories mostly
satisfy FAIR principles (eg: @Bonaretti2019-zs).
From the perspective of a researcher sharing genomic data, they are "good enough" as compared to the
well known alternatives (eg: general purpose repository, institutional data storage, supplementary file,
dropbox, on request, etc.).

@@

### R3 Point 22

> 22. Page 11, lines 13-15: That's not true for some of these disciplines. Ecology has NEON, evo bio
has NCBI.

RESPONSE: We acknowledge that NCBI is a well known repository for molecular evolution.
We have altered the section for clarity:

@@

> 23. Page 11, line 34: Need evidence to back this up.

RESPONSE: Reviewer 3 commented on the passage saying supplementary data files are less
findable and accessible as compared to FAIR data repositories.
We note that Roche and colleagues previously stated that [@Roche2015-aq]:

> There are no standards for organizing supplementary data both within and across journals
> and such data are often not readily discoverable or openly accessible
> (to those without a relevant journal subscription, for example).

They also cite an article by Caetano and Aisenberg that speaks directly to findability and
accessibility @Caetano2014-bi:

@@

> Projects involving computationally intensive tasks would benefit from a build/workflow automation
> solution.
> Guides for selecting and using such systems have been already reviewed by others [@Leipzig2016-bf;
> @Reiter2021-lh].
> But it is worth mentioning some of the most used solutions in bioinformatics, which include
> `snakemake` [@Koster2012-sq; @Molder2021-mp], `targets` [@Landau2021], `CWL` [@Crusoe2022-cb],
> `WDL` [@Voss2017-vd] and `nextflow` [@Di_Tommaso2017-yo].

### R3 Point 26

> 26. Page 14, lines 9-10: The big question is how to make more progress.
> We have the tools to achieve reproducibility, but why are we rarely achieving it?
> The paper mentions incentives and lack of training, which are true.
> You might consider elaborating a bit.
> My cynical view is that writing more papers and tutorials will do little without strong incentives
> and more automation.

RESPONSE: Thanks for raising this, and I note that similar sentiments were shared by reviewer 2 about the
challenges that are holding back progress.
Please refer to our response to Reviewer 2's Point #3.
The Challenges section now includes brief mentions of incentives and automation.

### 2023-08-28 manuscript/5pillars.Rmd
close #29 reviewer comments addressed now

@@

In software development, continuous integration/continuous delivery (CI/CD) is a process of automated
software compilation and testing.
A derivation of this for data analysis has been termed "continuous analysis", and involves automatic
execution and testing whenever a change is made to a project's code base
[@Beaulieu-Jones2017-vj; @Blischak2019-dq].
If the workflow executes successfully and tests are passed, the repository can be decorated with a
"badge" to inform users of the lack of errors in the code.
If those tests are designed cleverly such that certain conclusions drawn in the corresponding research
article can be assessed computationally, then such tests could be used to demonstrate computational
reproducibility without the need for individuals to execute the workflow themselves.
Continuous analysis might therefore be a future pediment (crowning feature) to the five-pillared
"temple" of reproducible research.

@@

Due to the relative lack of systematic reproducibility/replication studies in bioinformatics,
the severity of the reproducibility problem isn't well appreciated.
The studies that have been conducted point to reproducibility rates of ~10% in computational
life sciences [@Ioannidis2009-hv; @Zaringhalam2020-ms; @Samuel2022-ha], but further systematic studies
are required to bring attention to the issue.

@@

Another hindrance to reproducibility more broadly is the fixation on novel research directions rather
than meticulous confirmation of previous findings [@Garijo2013-zl ; @Huang2013-xr; @Kaelin2017-id;
@Munafo2017-ty; @Diong2021-is].
This is a direct result of research assessment policies that incentivize journal-level prestige
metrics and number of publications rather than the rigor [@Munafo2017-gf; @Oransky2023-zb].
A change in research assessment ("eg: track-record") that recognises and credits contributions toward
reproducibility and rigor would help decrease funds wasted on sloppy research.

>>>>

<hr>

@@

In software development, continuous integration/continuous delivery (CI/CD) is a process of automated
software compilation and testing.
A derivation of this for data analysis has been termed "continuous analysis", and involves automatic
execution and testing whenever a change is made to a project's code base
[@Beaulieu-Jones2017-vj; @Blischak2019-dq].
If the workflow executes successfully and tests are passed, the repository and linked research article
can be decorated with a "badge" to inform users of the lack of errors in the code,
similar to badges used for other Open Science practices [@Kidwell2016-bb].
If those tests are designed cleverly such that certain conclusions drawn in the corresponding research
article can be assessed computationally, then such tests could be used to demonstrate computational
reproducibility without the need for individuals to execute the workflow themselves [@Suetake2022-zq].

@@

Due to the relative lack of systematic reproducibility/replication studies in bioinformatics,
the severity of the reproducibility problem isn't well appreciated.
The studies that have been conducted point to reproducibility rates of ~10% in computational
life sciences [@Ioannidis2009-hv; @Zaringhalam2020-ms; @Samuel2022-ha], but further studies
are required to bring attention to the issue.

@@

Another hindrance to reproducibility more broadly is the fixation on novel research directions rather
than meticulous confirmation of previous findings [@Garijo2013-zl ; @Huang2013-xr; @Kaelin2017-id;
@Munafo2017-ty; @Diong2021-is].
This is a direct result of research assessment policies that incentivize journal-level prestige
metrics and number of publications rather than rigor [@Munafo2017-gf; @Oransky2023-zb].
A change in research assessment ("eg: track-record") that recognises and credits contributions toward
reproducibility and rigor would help decrease funds wasted on sloppy research.
